\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{authaffil}   % the package being documented (used live below)

% --- code listing style ---------------------------------------------------
\definecolor{codebg}{gray}{0.96}
\definecolor{codekw}{RGB}{30,80,160}
\lstdefinestyle{tex}{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{codebg},
  frame=single, framesep=6pt, rulecolor=\color{gray!40},
  breaklines=true, columns=fullflexible, keepspaces=true,
  xleftmargin=2pt, xrightmargin=2pt,
  language=[LaTeX]TeX,
  keywordstyle=\color{codekw},
  morekeywords={DefineAffiliation,AddAuthor,AddAffilNote,PrintAuthors,
    PrintAffiliations,PrintAffilNotes,usepackage},
  moredelim=[is][\color{codekw}\bfseries]{|}{|}
}
\lstset{style=tex}
\newcommand{\cs}[1]{\texttt{\textbackslash #1}}
\newcommand{\meta}[1]{$\langle$\textit{#1}$\rangle$}

\title{\textbf{The \texttt{authaffil} package}\\[2pt]
  \large First-appearance numbering of author affiliations}
\author{Ayush Noori}
\date{Version 1.1 \,\textperiodcentered\, 2026/05/29}

\begin{document}
\maketitle
\thispagestyle{empty}

\begin{abstract}
\noindent\texttt{authaffil} removes hand-numbered affiliation superscripts from
multi-author papers. You define each affiliation once under a short code, tag
each author with codes, and the package assigns and prints the numbers in the
order the codes are \emph{first cited}. Adding, removing, or reordering authors
renumbers everything automatically.
\end{abstract}

\section{Installation}
Place \texttt{authaffil.sty} next to your \texttt{.tex} file (or anywhere on
your \TeX\ path) and load it:
\begin{lstlisting}
\usepackage{authaffil}            % default: per-author numbers sorted ascending
\usepackage[nosort]{authaffil}    % per-author numbers in citation order
\end{lstlisting}
It requires only \texttt{expl3} (present in every modern \TeX\ distribution)
and works with pdf\LaTeX, Xe\LaTeX, and Lua\LaTeX.

\section{Quick start}
\begin{lstlisting}
% 1. Define affiliations (the order of these lines is irrelevant).
\DefineAffiliation{HMS}{Dept. of Biomedical Informatics, HMS, Boston, MA, USA}
\DefineAffiliation{OXE}{Dept. of Engineering Science, University of Oxford, UK}
\DefineAffiliation{ETH}{Dept. of Computer Science, ETH Zurich, Switzerland}

% 2. List authors, tagging each with codes (+ optional footnote symbols).
\AddAuthor{Ayush~Noori}{HMS,OXE}
\AddAuthor[*]{Manuel~Burger}{ETH}
\AddAuthor[\dagger,\ddagger]{Marinka~Zitnik}{HMS}

% 3. (Optional) legend lines.
\AddAffilNote{*}{Equal contribution.}
\AddAffilNote{\dagger}{Jointly supervised this work.}

% 4. Typeset -- e.g. inside \author{...} of your template.
\author{{\small\begin{center}
    \PrintAuthors      \\[2mm]
    \PrintAffiliations \\[1mm]
    \PrintAffilNotes
\end{center}}}
\end{lstlisting}

\noindent The block above renders as:

% --- executed setup matching the listing above, so the demo renders ---
\DefineAffiliation{HMS}{Dept. of Biomedical Informatics, HMS, Boston, MA, USA}
\DefineAffiliation{OXE}{Dept. of Engineering Science, University of Oxford, UK}
\DefineAffiliation{ETH}{Dept. of Computer Science, ETH Zurich, Switzerland}
\AddAuthor{Ayush~Noori}{HMS,OXE}
\AddAuthor[*]{Manuel~Burger}{ETH}
\AddAuthor[\dagger,\ddagger]{Marinka~Zitnik}{HMS}
\AddAffilNote{*}{Equal contribution.}
\AddAffilNote{\dagger}{Jointly supervised this work.}

\begin{center}\fbox{\parbox{0.92\textwidth}{\centering
  \PrintAuthors \\[2mm]
  \PrintAffiliations \\[1mm]
  \PrintAffilNotes
}}\end{center}

\section{Command reference}

\paragraph{\cs{DefineAffiliation}\texttt{\{}\meta{code}\texttt{\}\{}\meta{address}\texttt{\}}}
Register one affiliation. \meta{code} is any short string---three-letter codes
are convenient but not required. \meta{address} may contain \cs{\textbackslash}
for forced line breaks. Call once per affiliation, in any order.

\paragraph{\cs{AddAuthor}\texttt{[}\meta{symbols}\texttt{]\{}\meta{name}\texttt{\}\{}\meta{codes}\texttt{\}}}
Append one author. \meta{codes} is a comma-separated list of affiliation codes;
each new code receives the next number the first time it is seen. The optional
\meta{symbols} argument is extra superscript material (e.g.\ \texttt{*},
\cs{dagger}, \cs{ddagger}) appended after the numbers---write it pre-separated
the way you want it to read, e.g.\ \texttt{[\cs{dagger},\cs{ddagger}]}. Use
\texttt{\textasciitilde} inside names to keep them from breaking across lines.

\paragraph{\cs{AddAffilNote}\texttt{\{}\meta{symbol}\texttt{\}\{}\meta{text}\texttt{\}}}
Register one legend line, typeset as \textsuperscript{\meta{symbol}}\meta{text}. Lines print
in the order added.

\paragraph{\cs{PrintAuthors}}
Typeset author names, comma-separated, each with its computed superscript. Sets
no font size---wrap it yourself, e.g.\ \texttt{\{\cs{small} \cs{PrintAuthors}\}}.

\paragraph{\cs{PrintAffiliations}}
Typeset the numbered affiliation list, one per line, in first-appearance order.
Switches to \cs{footnotesize}.

\paragraph{\cs{PrintAffilNotes}}
Typeset the legend lines, one per line. Switches to \cs{footnotesize}.

\section{Options}
\begin{description}[leftmargin=2.2cm,style=nextline,font=\ttfamily]
  \item[\lbrack sort\rbrack] (default) Numbers within a single author are shown
    ascending: \texttt{1,5,30}, regardless of citation order.
  \item[\lbrack nosort\rbrack] Numbers within a single author are shown in the
    order the codes were cited.
\end{description}

\section{How the numbering works}
A code receives its number the first time \emph{any} author cites it. So for an
author who introduces several brand-new affiliations at once, the left-to-right
order of that author's code list fixes those numbers. To force a particular
global ordering, control the order of first citation---the simplest lever is the
first author who uses each code. Everything downstream (the per-author
superscripts and the printed affiliation list) is derived from that, so there
are never two places to keep in sync.

\section{Diagnostics}
The package writes a warning to the log and terminal in two situations:
\begin{description}[leftmargin=2.6cm,style=nextline,font=\ttfamily]
  \item[unknown-code] An author cites a code that was never defined with
    \cs{DefineAffiliation}. A number is still allocated, but no address exists
    for it.
  \item[unused-affil] A code was defined but never cited by any author, so it
    does not appear in the list. Checked once, at \cs{end\{document\}}.
\end{description}
Neither is fatal; both are there to catch copy-paste slips before submission.

\section{Tips}
\begin{itemize}[leftmargin=1.4em]
  \item The commands are layout-agnostic. Drop \cs{PrintAuthors},
    \cs{PrintAffiliations}, and \cs{PrintAffilNotes} into whatever
    \cs{title}/\cs{author} scaffold your journal class expects.
  \item A trailing \cs{\textbackslash} immediately before \cs{end\{center\}}
    raises a spurious ``missing \cs{item}'' error in \LaTeX. The print commands
    therefore use \cs{\textbackslash} as a \emph{separator}, never a terminator,
    so you can place them safely at the end of a centered block.
  \item If a legend line uses \cs{href}, load \texttt{hyperref} in your
    preamble; the package does not load it for you.
\end{itemize}

\end{document}
